Examine the efficacy of ascaroside #18 in controling Escherichia Coli on alfalfa and fenugreek seeds and sprouts

### 2024-02-23 R/eda-code/eda-v2.qmd
part 2 first time
about to push but found some problems so just push the changes I have firstly

---
title: "An example exploratory analysis script"
author: "Andreas Handel"
date: "10/28/2021"
output: html_document
---

@@

I'm only showing it the way where the code is included in the file. 
As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.

But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)

Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.

@@

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
```

@@

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

@@

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

@@

```{r}
# Fit the ANOVA model
model <- aov(Selective ~ Seed + Strain + Treatment + Day + Rep, data = mydata)

# Summarize the ANOVA results
linear_model <- summary(model)

# Save this table

anovatable_file = here("results", "tables", "anovatable.rds")
saveRDS(linear_model, file = anovatable_file)
```

>>>>

---
title: "Project data cleaning"
author: "Xueyan Hu"
date: "2023-02-22"
output: html_document
---

@@

This is essentially the same as the other `exploratory_analysis` Quarto file, but now the code is not inside this file. Instead, it is pulled in from the R script `exploratorycode.R` using the code chunk labels.

@@

# Setup
Load the chunk.

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('edacode.R')
```

Load the packages.
```{r,packages, echo=FALSE,message=FALSE}
```

@@

```{r,loaddata}
```

@@

```{r,table1}
```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# Data exploration through figures

Seed type first: bar chart

```{r,seed}
```

Then goes strain type: box plot

```{r,strain}
```

Now the effect of the treatment: bar chart with color gradient as an indicator

```{r,treatment}
```

The growth trend of the bacterial population over time: scatterplot with line

```{r,day}
```

@@

```{r,modelfit}
```

### 2024-02-23 R/eda-code/eda.qmd
part 2 first time
about to push but found some problems so just push the changes I have firstly

linear model applied

```{r}
# Fit the ANOVA model
model <- aov(Selective ~ Seed + Strain + Treatment + Day + Rep, data = mydata)

# Summarize the ANOVA results
linear_model <- summary(model)

# Save this table

anovatable_file = here("results", "tables", "anovatable.rds")
saveRDS(linear_model, file = anovatable_file)
```

>>>>

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# Data exploration through figures

Seed type first: bar chart

```{r}
# Calculate average population by seed type
mydata_seed <- mydata %>%
  group_by(Seed) %>%
  summarize(mean_Selective = mean(Selective),
            sd_Selective = sd(Selective))

# Create the bar chart with standard deviation lines
bp1 <- ggplot(mydata_seed, aes(x = Seed, y = mean_Selective, fill = Seed)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7, width = 0.4)  +
  geom_errorbar(data = mydata_seed, aes(x = Seed, ymin = mean_Selective - sd_Selective, ymax = mean_Selective + sd_Selective, color = "black"),
                position = position_dodge(width = 0.1), width = 0.05) +
  labs(title = "The population difference by seed type",
       x = "Seed Type",
       y = "Average Population") +
  scale_color_manual(values = c("Alfalfa" = "lightblue", "Fenugreek" = "pink")) +
  theme_minimal()
plot(bp1)
figure_file = here("results","figures", "seed_type.png")
ggsave(filename = figure_file, plot=bp1) 
```

Then goes strain type: box plot

```{r}
# Create boxplot for bacterial population by 
bp2 <- ggplot(mydata, aes(x = Strain, y = Selective)) +
  geom_boxplot() +
  labs(title = "The population difference by treatment",
       x = "Strain type",
       y = "Average Population") +
  theme_minimal()
plot(bp2)
figure_file = here("results","figures", "strain_type.png")
ggsave(filename = figure_file, plot=bp2) 
```

Now the effect of the treatment: bar chart with color gradient as an indicator

```{r}
# Calculate average population by seed type
mydata_treatment <- mydata %>%
  group_by(Treatment) %>%
  summarize(mean_Selective = mean(Selective),
            sd_Selective = sd(Selective))

# Create a continuous variable for fill
mydata_treatment$fill_color <- scales::rescale(mydata_treatment$mean_Selective)

# Create the bar chart with standard deviation lines
bp3 <- ggplot(mydata_treatment, aes(x = Treatment, y = mean_Selective, fill = fill_color)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7, width = 0.4)  +
  geom_errorbar(data = mydata_treatment, aes(x = Treatment, ymin = mean_Selective - sd_Selective, ymax = mean_Selective + sd_Selective, color = "grey"),
                position = position_dodge(width = 0.1), width = 0.05) +
  labs(title = "The population difference by treatment",
       x = "Treatment",
       y = "Average Population") +
  scale_fill_gradient(low = "yellow", high = "red") +  # Adjust the colors as needed
  scale_color_manual(values = c("grey")) +  # Use grey color for error bar color
  guides(color = "none") +  # Turn off both fill and color legends
  theme_minimal()
plot(bp3)
figure_file = here("results","figures", "treatment.png")
ggsave(filename = figure_file, plot=bp3) 
```

The growth trend of the bacterial population over time: scatterplot with line

```{r}
# Calculate the average population for each day
mydata_day <- aggregate(Selective ~ Day, data = mydata, FUN = mean)

# Draw a point plot
sp4 <- ggplot(mydata_day, aes(x = Day, y = Selective)) +
  geom_point(size = 2, color = "blue") +
  geom_line(aes(group = 1), color = "red") +
  labs(title = "The growth trend of the bacterial population over time",
       x = "Day",
       y = "Average Bacterial Population") +
  theme_minimal()
plot(sp4)
figure_file = here("results","figures", "timepoint.png")
ggsave(filename = figure_file, plot=sp4) 
```
```

### 2024-02-23 R/processing-code/processingfile-v2.qmd
part 2 first time
about to push but found some problems so just push the changes I have firstly

# Processing script

This contains the same code and comments/information as `processingcode.R`.

This just shows it as Quarto file, to give you an idea how to do it in a setup that combines code and text in a single file.

See the other Quarto file for my currently preferred approach of pulling code from the R script into the Quarto file.

@@

Load needed packages. make sure they are installed.

```{r}
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
```

@@

```{r}
# path to data
# note the use of the here() package and not absolute paths
data_location <- here::here("data","raw-data","raw_data.xlsx")
rawdata <- readxl::read_excel(data_location)
```

# Check data

First we can look at the codebook

```{r}
codebook <- readxl::read_excel(data_location, sheet ="Codebook")
print(codebook)
```

@@

```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```

@@

```{r}
data1 <- rawdata %>%
  mutate(Day = paste0("Day", substring(Day, 2)))
skimr::skim(data1)
head(data1)
```

@@

@@

```{r}
processeddata <- data1
save_data_location <- here::here("data","processed-data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```

# Notes

Removing anyone who had "faulty" or missing data is one approach. It's often not the best. based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).

>>>>

# Processing script with code pulled in

This is essentially the same as the other Quarto file, but now the code is not inside this file. Instead, it is pulled in from the R script using the code chunk labels.

@@

This needs to run to load the R script so we can include the code chunks below.

Note that you can pull in more than one R script if you want.

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('processingcode.R')
```

Load needed packages. Instead of having R commands here, this just pulls in the code from the R script we loaded above and labelled.

```{r, packages, message = FALSE, warning = FALSE}
```

@@

```{r, loaddata}
```

# Explore data

@@

```{r, exploredata}
```

@@

```{r, cleandata1}
```

@@

All done, data is clean now. Time to save.
Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

@@

```{r, savedata}
```

### 2024-02-23 products/manuscript/Manuscript copy.qmd
part 2 first time
about to push but found some problems so just push the changes I have firstly

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Xueyan HU
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: dataanalysis-references.bib
csl: apa.csl
editor: 
  markdown: 
    wrap: sentence
---

@@

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key.
Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

@@

*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*

@@

*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*

@@

>>>>

---
title: "Examine the efficacy of ascaroside #18 in controling Escherichia Coli on alfalfa and fenugreek seeds and sprouts"
subtitle: ""
author: Xueyan Hu
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: dataanalysis-references.bib
csl: apa.csl
editor: 
  markdown: 
    wrap: sentence
---

@@

@@

Since the dataset is quite well-organized and no any value missing, the cleaning step is relatively easy.
I only changed the category names.

My main interest is to test if the chemical treatment is effective or not compared to the control treated groups.
Also, anova model will be applied to check which variable will make significant different to the bacterial population.
So boxplot, bar chart and line chart will be created according to the analysis direction.

@@

All of the data in this dataset was collected by me in 2022 after experiment was conducted.

@@

The detailed raw data and processed data are stored in 'data' folder.
